---
title: "Scrape Games"
date: "07-26-2022"
author: "Luke Welsh"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,warning=F,message=F)
library(baseballr)
library(dplyr)
library(tidyverse)
library(janitor)
library(ggplot2)
library(readxl)
```

In this Document, I will:

1. Create functions to scrape MLB pitch-by-pitch data from baseball-savant.com, summarize games, and join all game summaries together.

2. Use these functions together to produce one large table with average exit velocity for each game.

3. Add additional statistic which will correct for certain teams simply having better batters and thus higher exit velocities on average. 

4. Join MLB data with TV broadcast data to add column with which station each game was broadcast to.


Example of Pitch-by-Pitch Game Data.
```{r}
pitches = get_pbp_mlb(663046) %>% 
  drop_na(hitData.launchSpeed) %>%
  select(matchup.batter.fullName, batting_team, hitData.launchSpeed)
head(pitches)
```

Fuction to break down full game stats into average exit velo by team.
```{r}
sum_game = function(game_pk){
  sum_table = get_pbp_mlb(game_pk) %>% 
  drop_na(hitData.launchSpeed) %>% 
  filter(hitData.trajectory != 'bunt_grounder') %>% 
  select(matchup.batter.fullName, batting_team, hitData.launchSpeed, game_date) %>% 
  group_by(batting_team) %>% 
  summarize(velo = mean(hitData.launchSpeed), game = game_pk, date = first(game_date))
  sum_table
}
sum_game(663046)
```

Function to combine each team's average exit velo from each game into one large tibble.
```{r}
join_games = function(game_ids){
  all_games = sum_game(game_ids[1])
  for (i in 2:length(game_ids)){
    all_games = rbind(all_games, sum_game(game_ids[i]))
    if(i %% 10 == 0){
      print(i)
    }
  }
  all_games
}
```

Getting each game's game_pk (id number).

Filters are to eliminate preseason games, games going on while the code runs, and 'DR'.

'DR' means a game that was postponed.
```{r}
schedule = mlb_schedule(2022) %>% 
  filter(series_description == "Regular Season") %>% 
  filter(status_abstract_game_state == "Final") %>% 
  filter(status_status_code != 'DR')
completed_games_pk = schedule %>% 
  pull(game_pk)
```

**TAKES A LONG TIME (~1 hour)**
Using previously written function to join information from every game.

Save resulting tibble as a csv so I don't have to re-run the big function between sessions.
```{r}
# all_games = join_games(completed_games_pk)
# write.csv(all_games, "all_game_sum.csv")

all_games = read.csv("all_game_sum.csv")
```

Adding 2 additional columns.

  First, team_avg is the corresponding team's average game exit velocity over the season.

  Secondly, game_diff is a statistic that finds the difference the team had from their team_avg in a specific game. This is because different teams are, on average, better than others in terms of average exit velocity. By finding the difference in a single game from their season average, the column game_diff is independent of the batting team and, assuming balls do not travel faster when games are on various broadcasts, the variation of this statistic should be just due to randomness.
```{r}
all_games = all_games %>% 
  group_by(batting_team) %>% 
  mutate(team_avg = mean(velo)) %>% 
  ungroup() %>% 
  mutate(game_diff = velo - team_avg)
```

Importing independently gathered National TV Schedule
```{r}
tv_schedule = read_excel("../juicy/national_tv_list.xlsx") %>% 
  rename(game = game_pk) %>% 
  select('Team 1', 'Team 2', 'TV Provider', 'game')
```

Joining together MLB games with TV schedule. 

In essence, adding a column with where each game was broadcast.
```{r}
all_games_tv = all_games %>% left_join(tv_schedule, by = 'game', keep = FALSE) %>% 
  select(-'Team 1', -'Team 2')
all_games_tv[is.na(all_games_tv)] = 'Local'
all_games_tv = all_games_tv %>%
  rename(tv_provider = `TV Provider`)
```

Saving data, this time with TV broadcast information added in.
```{r}
write.csv(all_games_tv, "all_game_tv.csv")
```















